---
title: "Lucas_Massier_Cohort2"
output: html_notebook
---
```{r}
set.seed(001)
```

```{r, results='hide'}
library(dplyr)
library(Seurat)
library(patchwork)
library(DoubletFinder)
```

```{r}
setwd("//home/lucamannino/R_Projects/scAtlas/Individual_Studies/LucasMassier")
```


```{r, results='hide'}
# Load the PBMC dataset
Patient21L005043 <- Read10X(data.dir = "./21L005043/")
Patient21L005043
# Initialize the Seurat object with the raw (non-normalized data).
Patient21L005043 <- CreateSeuratObject(counts = Patient21L005043, project = "Patient21L005043", min.cells = 3, min.features = 200)
Patient21L005043
```
36601 genes 4606 samples obtained from cell ranger
After removing cell with less than 200 gene counts and removing genes expressed in less than 3 cells:
25861 features across 4606 samples within 1 assay 


```{r, results='hide'}
# Load the PBMC dataset
Patient21L005039 <- Read10X(data.dir = "./21L005039/")
Patient21L005039
# Initialize the Seurat object with the raw (non-normalized data).
Patient21L005039 <- CreateSeuratObject(counts = Patient21L005039, project = "Patient21L005039", min.cells = 3, min.features = 200)
Patient21L005039
```
36601 genes  1829 samples obtained from cell ranger
After removing cell with less than 200 gene counts and removing genes expressed in less than 3 cells:
21958 features across 1829  samples within 1 assay 






```{r}
GSE225700 <- c(Patient21L005039,Patient21L005043)
names(GSE225700) <- c("Patient21L005039","Patient21L005043")
```

```{r}


GSE225700$Patient21L005039[["mt.percent"]] <-PercentageFeatureSet(GSE225700$Patient21L005039, pattern = "^MT-")
GSE225700$Patient21L005043[["mt.percent"]] <-PercentageFeatureSet(GSE225700$Patient21L005043, pattern = "^MT-")





```
```{r}
lapply(GSE225700, function(x)  {VlnPlot(x, features = c("nFeature_RNA", "nCount_RNA", "mt.percent"), ncol = 3)})
```
2 single cell 2 single nuclei?

```{r}





max.mito.thr_Patient21L005039 <- median(GSE225700$Patient21L005039@meta.data$mt.percent) + 4*mad(GSE225700$Patient21L005039@meta.data$mt.percent)
max.mito.thr_Patient21L005043 <- median(GSE225700$Patient21L005043@meta.data$mt.percent) + 4*mad(GSE225700$Patient21L005043@meta.data$mt.percent)




max.mito.thr_Patient21L005039
max.mito.thr_Patient21L005043


#For Single nuclei wont be using MT cutoff as dying cells shouldn't be noticeable by MT percentage
```




```{r}
library(ggExtra)
library(ggplot2)
p2 <- ggplot(GSE225700$Patient21L005039@meta.data, aes(x=nFeature_RNA, y=mt.percent)) +
      geom_point() +
      geom_hline(aes(yintercept = max.mito.thr_Patient21L005039), colour = "red", linetype = 2) #+
#      annotate(geom = "text", label = paste0(as.numeric(table(GSE225700$Sample1@meta.data$mt.percent > max.mito.thr_SRR12423009)))," cells removed\n",                                        as.numeric(table(GSE225700$Sample1@meta.data$mt.percent > max.mito.thr_SRR12423009)," cells remain"), x = 6000, y = 0.1)

ggMarginal(p2, type = "histogram", fill="lightgrey", bins=100) 
```

```{r}
library(ggExtra)
library(ggplot2)
p3 <- ggplot(GSE225700$Patient21L005043@meta.data, aes(x=nFeature_RNA, y=mt.percent)) +
      geom_point() +
      geom_hline(aes(yintercept = max.mito.thr_Patient21L005043), colour = "red", linetype = 2) #+
#      annotate(geom = "text", label = paste0(as.numeric(table(GSE225700$Sample1@meta.data$mt.percent > max.mito.thr_SRR12423009)))," cells removed\n",                                        as.numeric(table(GSE225700$Sample1@meta.data$mt.percent > max.mito.thr_SRR12423009)," cells remain"), x = 6000, y = 0.1)

ggMarginal(p3, type = "histogram", fill="lightgrey", bins=100) 
```




```{r}

lapply(GSE225700, function(x)  {FeatureScatter(x, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")})
lapply(GSE225700, function(x)  {ggMarginal(ggplot(x@meta.data, aes(x=nFeature_RNA, y=nCount_RNA))+ geom_point() , type = "histogram", fill="lightgrey", bins=100)})
```

```{r, results='hide'}

GSE225700$Patient21L005039
GSE225700$Patient21L005043



```

Patient21L005039:
21958 features across 1829 samples within 1 assay 
Active assay: RNA (21958 features, 0 variable features)
Patient21L005043:
25861 features across 4606 samples within 1 assay 
Active assay: RNA (25861 features, 0 variable features)



```{r}



GSE225700$Patient21L005039 <- subset(GSE225700$Patient21L005039, subset = nFeature_RNA < 6500  & nCount_RNA < 40000 )

GSE225700$Patient21L005043 <- subset(GSE225700$Patient21L005043, subset = nFeature_RNA < 7500   & nCount_RNA < 40000 )




```






```{r, results='hide'}

GSE225700$Patient21L005039
GSE225700$Patient21L005043



```


Patient21L005039
21958 features across 1820 samples within 1 assay 
Active assay: RNA (21958 features, 0 variable features)
Patient21L005043
25861 features across 4572 samples within 1 assay 
Active assay: RNA (25861 features, 0 variable features)



I will keep going with strict version:



```{r}

lin_mod <- lm(log(GSE225700$Patient21L005039@meta.data$nFeature_RNA) ~ log(GSE225700$Patient21L005039@meta.data$nCount_RNA))
sub <- lin_mod$model[lin_mod$residuals < -0.5,]
plot(log(GSE225700$Patient21L005039@meta.data$nCount_RNA),log(GSE225700$Patient21L005039@meta.data$nFeature_RNA), xlab ="log(nCount_RNA)",ylab="log(nFeature_RNA)",pch=20,col="grey") +
abline(lin_mod,col="red",lwd=3) +
points(sub[,2], sub[,1],pch=20,col="blue")
tokeep <- as.numeric(names(lin_mod$residuals[lin_mod$residuals >= -0.5]))
GSE225700$Patient21L005039 <- subset(GSE225700$Patient21L005039, cells = tokeep)
```


```{r}

lin_mod <- lm(log(GSE225700$Patient21L005043@meta.data$nFeature_RNA) ~ log(GSE225700$Patient21L005043@meta.data$nCount_RNA))
sub <- lin_mod$model[lin_mod$residuals < -0.5,]
plot(log(GSE225700$Patient21L005043@meta.data$nCount_RNA),log(GSE225700$Patient21L005043@meta.data$nFeature_RNA), xlab ="log(nCount_RNA)",ylab="log(nFeature_RNA)",pch=20,col="grey") +
abline(lin_mod,col="red",lwd=3) +
points(sub[,2], sub[,1],pch=20,col="blue")
tokeep <- as.numeric(names(lin_mod$residuals[lin_mod$residuals >= -0.5]))
GSE225700$Patient21L005043 <- subset(GSE225700$Patient21L005043, cells = tokeep)
```





```{r}
Patient21L005039DoubletPerc <- 0.008*(1829/1000)
Patient21L005039DoubletPerc
```
DB ratio 0.014632

```{r}
GSE225700$Patient21L005039 <- GSE225700$Patient21L005039 %>% NormalizeData(normalization.method = "LogNormalize", scale.factor = 10000) %>% FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% ScaleData(vars.to.regress = c("nCount_RNA", "nFeature_RNA")) %>% RunPCA(npcs = 100)

ElbowPlot(GSE225700$Patient21L005039, ndims = 50)

sweep.res.list_GSE225700Patient21L005039 <- paramSweep_v3(GSE225700$Patient21L005039, PCs = 1:30, sct = FALSE)
sweep.stats_GSE225700Patient21L005039 <- summarizeSweep(sweep.res.list_GSE225700Patient21L005039, GT = FALSE)
bcmvn_GSE225700Patient21L005039 <- find.pK(sweep.stats_GSE225700Patient21L005039)

GSE225700$Patient21L005039 <- FindNeighbors(GSE225700$Patient21L005039, dims = 1:30, reduction = "pca", features = VariableFeatures(GSE225700$Patient21L005039))
GSE225700$Patient21L005039 <- FindClusters(GSE225700$Patient21L005039)


homotypic.prop_GSE225700Patient21L005039 <- modelHomotypic(GSE225700$Patient21L005039@meta.data$seurat_clusters)   
nExp_poi_GSE225700Patient21L005039 <- round(Patient21L005039DoubletPerc*nrow(GSE225700$Patient21L005039@meta.data))
nExp_poi.adj_GSE225700Patient21L005039 <- round(nExp_poi_GSE225700Patient21L005039*(1-homotypic.prop_GSE225700Patient21L005039))


pK_GSE225700Patient21L005039 = as.numeric(as.character(bcmvn_GSE225700Patient21L005039$pK))
BCmetric_GSE225700Patient21L005039 = bcmvn_GSE225700Patient21L005039$BCmetric
pK_choose_GSE225700Patient21L005039 = pK_GSE225700Patient21L005039[which(BCmetric_GSE225700Patient21L005039 %in% max(BCmetric_GSE225700Patient21L005039))]

par(mar=c(5,4,4,8)+1,cex.main=1.2,font.main=2)
plot(x = pK_GSE225700Patient21L005039, y = BCmetric_GSE225700Patient21L005039, pch = 16,type="b",
col = "blue",lty=1)
abline(v=pK_choose_GSE225700Patient21L005039,lwd=2,col='red',lty=2)
title("The BCmvn distributions")
text(pK_choose_GSE225700Patient21L005039,max(BCmetric_GSE225700Patient21L005039),as.character(pK_choose_GSE225700Patient21L005039),pos = 4,col = "red")
```

use appropiate pK value as estimated above
```{r}
GSE225700$Patient21L005039 <- doubletFinder_v3(GSE225700$Patient21L005039, PCs = 1:30, pN = 0.25, pK = pK_choose_GSE225700Patient21L005039, nExp = nExp_poi.adj_GSE225700Patient21L005039, reuse.pANN = FALSE, sct = FALSE)

#
```
```{r}
GSE225700$Patient21L005039 <- RunUMAP(GSE225700$Patient21L005039, dims = 1:20,reduction = "pca", verbose = FALSE)
```

```{r}
DimPlot(GSE225700$Patient21L005039,group.by="DF.classifications_0.25_0.19_24")
```




```{r}
GSE225700$Patient21L005039
GSE225700$Patient21L005039 <- subset(GSE225700$Patient21L005039, subset = DF.classifications_0.25_0.19_24 == "Singlet")
GSE225700$Patient21L005039 
```
From: 21958 features across 1820 samples

to:
21958 features across 1796 samples 

```{r}
Patient21L005043DoubletPerc <- 0.008*(4606/1000)
Patient21L005043DoubletPerc
```
DB ratio 0.036848

```{r}
GSE225700$Patient21L005043 <- GSE225700$Patient21L005043 %>% NormalizeData(normalization.method = "LogNormalize", scale.factor = 10000) %>% FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% ScaleData(vars.to.regress = c("nCount_RNA", "nFeature_RNA")) %>% RunPCA(npcs = 100)

ElbowPlot(GSE225700$Patient21L005043, ndims = 50)

sweep.res.list_GSE225700Patient21L005043 <- paramSweep_v3(GSE225700$Patient21L005043, PCs = 1:30, sct = FALSE)
sweep.stats_GSE225700Patient21L005043 <- summarizeSweep(sweep.res.list_GSE225700Patient21L005043, GT = FALSE)
bcmvn_GSE225700Patient21L005043 <- find.pK(sweep.stats_GSE225700Patient21L005043)

GSE225700$Patient21L005043 <- FindNeighbors(GSE225700$Patient21L005043, dims = 1:30, reduction = "pca", features = VariableFeatures(GSE225700$Patient21L005043))
GSE225700$Patient21L005043 <- FindClusters(GSE225700$Patient21L005043)


homotypic.prop_GSE225700Patient21L005043 <- modelHomotypic(GSE225700$Patient21L005043@meta.data$seurat_clusters)   
nExp_poi_GSE225700Patient21L005043 <- round(Patient21L005043DoubletPerc*nrow(GSE225700$Patient21L005043@meta.data))
nExp_poi.adj_GSE225700Patient21L005043 <- round(nExp_poi_GSE225700Patient21L005043*(1-homotypic.prop_GSE225700Patient21L005043))


pK_GSE225700Patient21L005043 = as.numeric(as.character(bcmvn_GSE225700Patient21L005043$pK))
BCmetric_GSE225700Patient21L005043 = bcmvn_GSE225700Patient21L005043$BCmetric
pK_choose_GSE225700Patient21L005043 = pK_GSE225700Patient21L005043[which(BCmetric_GSE225700Patient21L005043 %in% max(BCmetric_GSE225700Patient21L005043))]

par(mar=c(5,4,4,8)+1,cex.main=1.2,font.main=2)
plot(x = pK_GSE225700Patient21L005043, y = BCmetric_GSE225700Patient21L005043, pch = 16,type="b",
col = "blue",lty=1)
abline(v=pK_choose_GSE225700Patient21L005043,lwd=2,col='red',lty=2)
title("The BCmvn distributions")
text(pK_choose_GSE225700Patient21L005043,max(BCmetric_GSE225700Patient21L005043),as.character(pK_choose_GSE225700Patient21L005043),pos = 4,col = "red")
```

use appropiate pK value as estimated above
```{r}
GSE225700$Patient21L005043 <- doubletFinder_v3(GSE225700$Patient21L005043, PCs = 1:30, pN = 0.25, pK = pK_choose_GSE225700Patient21L005043, nExp = nExp_poi.adj_GSE225700Patient21L005043, reuse.pANN = FALSE, sct = FALSE)

#
```
```{r}
GSE225700$Patient21L005043 <- RunUMAP(GSE225700$Patient21L005043, dims = 1:20,reduction = "pca", verbose = FALSE)
```

```{r}
DimPlot(GSE225700$Patient21L005043,group.by="DF.classifications_0.25_0.06_153")
```




```{r}
GSE225700$Patient21L005043
GSE225700$Patient21L005043 <- subset(GSE225700$Patient21L005043, subset = DF.classifications_0.25_0.06_153 == "Singlet")
GSE225700$Patient21L005043 
```
from:25861 features across 4572 samples
to: 25861 features across 4419 samples
21L005043 is diabetic
21L005039 healthy obese
sex:female


```{r}
LucasMassierCohort2 <- merge(GSE225700$Patient21L005043,GSE225700$Patient21L005039)
```


```{r}
LucasMassierCohort2$sex <- "female"
LucasMassierCohort2$Condition <- as.character(LucasMassierCohort2$orig.ident)
LucasMassierCohort2$Condition  <- replace(LucasMassierCohort2$Condition, LucasMassierCohort2$Condition %in% "Patient21L005043", "Healthy Obese")
LucasMassierCohort2$Condition  <- replace(LucasMassierCohort2$Condition, LucasMassierCohort2$Condition %in% "Patient21L005039", "Diabetic Obese")
LucasMassierCohort2$technology<- "Chromium-v3 single nuclei"
LucasMassierCohort2$Study_technology <- "Luca Massier Chromium-v3 single nuclei"
LucasMassierCohort2$Study <- "Luca Massier"
LucasMassierCohort2$Nuclei_Cell <- "SingleNuclei"
```


```{r}
saveRDS(LucasMassierCohort2,"LucasMassierCohort2.RDS")
```


