---
title: "New_GSE128889"
output: html_notebook
---


```{r}
set.seed(001)
```

```{r, results='hide'}
library(dplyr)
library(Seurat)
library(patchwork)
library(DoubletFinder)
```

```{r}
setwd("/home/lucamannino/R_Projects/scAtlas/Individual_Studies/GSE128889")
```

```{r}

# Load the PBMC dataset
GSE128889 <- Read10X(data.dir = "./GSE128889_RAW")
# Initialize the Seurat object with the raw (non-normalized data).
GSE128889
GSE128889 <- CreateSeuratObject(counts = GSE128889, project = "GSE128889", min.cells = 3, min.features = 200)
GSE128889
```
All data obtained from cellRanger
33538 genes across 11327 samples
After cells and genes with low counts are removed
17293 genes across 11325 samples


```{r}
GSE128889[["mt.percent"]] <-PercentageFeatureSet(GSE128889, pattern = "^MT-")

VlnPlot(GSE128889, features = c("nFeature_RNA", "nCount_RNA", "mt.percent"), ncol = 3)
```

Proceed to filter cells


```{r}
max.mito.thr_GSE128889 <- median(GSE128889@meta.data$mt.percent) + 4*mad(GSE128889@meta.data$mt.percent)
```


```{r}
library(ggExtra)
library(ggplot2)
p1 <- ggplot(GSE128889@meta.data, aes(x=nFeature_RNA, y=mt.percent)) +
      geom_point() +
      geom_hline(aes(yintercept = max.mito.thr_GSE128889), colour = "red", linetype = 2) #+
#      annotate(geom = "text", label = paste0(as.numeric(table(GSE128889_Filtered@meta.data$mt.percent > max.mito.thr_GSE128889_Filtered)))," cells removed\n",                                        as.numeric(table(GSE128889_Filtered@meta.data$mt.percent > max.mito.thr_GSE128889_Filtered)," cells remain"), x = 6000, y = 0.1)

ggMarginal(p1, type = "histogram", fill="lightgrey", bins=500) 
```

```{r}
ggMarginal(ggplot(GSE128889@meta.data, aes(x=nFeature_RNA, y=nCount_RNA))+ geom_point() , type = "histogram", fill="lightgrey", bins=100)
FeatureScatter(GSE128889, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
ggMarginal(ggplot(GSE128889@meta.data, aes(x=nFeature_RNA, y=nCount_RNA))+ geom_point() , type = "histogram", fill="lightgrey", bins=100)
```




 proceed to filter out low quality cells
```{r}
GSE128889
GSE128889_Filtered <- subset(GSE128889, subset = nFeature_RNA < 3000  & mt.percent < max.mito.thr_GSE128889 & nCount_RNA < 12500 )
GSE128889_Filtered
```

from 18433 features across 11325 samples
to: 18433 features across 10915 samples



check linear model after filtering
```{r}
lin_mod <- lm(log(GSE128889_Filtered@meta.data$nFeature_RNA) ~ log(GSE128889_Filtered@meta.data$nCount_RNA))
sub <- lin_mod$model[lin_mod$residuals < -0.5,]
plot(log(GSE128889_Filtered@meta.data$nCount_RNA),log(GSE128889_Filtered@meta.data$nFeature_RNA), xlab ="log(nCount_RNA)",ylab="log(nFeature_RNA)",pch=20,col="grey") +
abline(lin_mod,col="red",lwd=3) +
points(sub[,2], sub[,1],pch=20,col="blue")
tokeep <- as.numeric(names(lin_mod$residuals[lin_mod$residuals >= -0.5]))
GSE128889_Filtered <- subset(GSE128889_Filtered, cells = tokeep)
```

```{r}
GSE128889_Filtered
```
18433 features across 10913  cells





Nest step remove doublets with doublet finder
estimated the number of doublets by multiplying the number of total cells sequenced by 0.08 per 1000 cells
First step is estimate percentage of doublets:
```{r}
GSE128889_FilteredDoubletPerc <- 0.008*(11327/1000)
GSE128889_FilteredDoubletPerc
```
DB proportion: 0.090616

```{r}
GSE128889_Filtered <- GSE128889_Filtered %>% NormalizeData(normalization.method = "LogNormalize", scale.factor = 10000) %>% FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% ScaleData(vars.to.regress = c("nCount_RNA", "nFeature_RNA")) %>% RunPCA(npcs = 100)

ElbowPlot(GSE128889_Filtered, ndims = 50)

sweep.res.list_GSE128889_Filtered <- paramSweep_v3(GSE128889_Filtered, PCs = 1:30, sct = FALSE)
sweep.stats_GSE128889_Filtered <- summarizeSweep(sweep.res.list_GSE128889_Filtered, GT = FALSE)
bcmvn_GSE128889_Filtered <- find.pK(sweep.stats_GSE128889_Filtered)

GSE128889_Filtered <- FindNeighbors(GSE128889_Filtered, dims = 1:30, reduction = "pca", features = VariableFeatures(GSE128889_Filtered))
GSE128889_Filtered <- FindClusters(GSE128889_Filtered)


homotypic.prop_GSE128889_Filtered <- modelHomotypic(GSE128889_Filtered@meta.data$seurat_clusters)   
nExp_poi_GSE128889_Filtered <- round(GSE128889_FilteredDoubletPerc*nrow(GSE128889_Filtered@meta.data))
nExp_poi.adj_GSE128889_Filtered <- round(nExp_poi_GSE128889_Filtered*(1-homotypic.prop_GSE128889_Filtered))


pK_GSE128889_Filtered = as.numeric(as.character(bcmvn_GSE128889_Filtered$pK))
BCmetric_GSE128889_Filtered = bcmvn_GSE128889_Filtered$BCmetric
pK_choose_GSE128889_Filtered = pK_GSE128889_Filtered[which(BCmetric_GSE128889_Filtered %in% max(BCmetric_GSE128889_Filtered))]

par(mar=c(5,4,4,8)+1,cex.main=1.2,font.main=2)
plot(x = pK_GSE128889_Filtered, y = BCmetric_GSE128889_Filtered, pch = 16,type="b",
col = "blue",lty=1)
abline(v=pK_choose_GSE128889_Filtered,lwd=2,col='red',lty=2)
title("The BCmvn distributions")
text(pK_choose_GSE128889_Filtered,max(BCmetric_GSE128889_Filtered),as.character(pK_choose_GSE128889_Filtered),pos = 4,col = "red")
```

use appropiate pK value as estimated above
```{r}
GSE128889_Filtered <- doubletFinder_v3(GSE128889_Filtered, PCs = 1:30, pN = 0.25, pK = pK_choose_GSE128889_Filtered, nExp = nExp_poi.adj_GSE128889_Filtered, reuse.pANN = FALSE, sct = FALSE)
GSE128889_Filtered
#
```

```{r}
GSE128889_Filtered <- RunUMAP(GSE128889_Filtered, dims = 1:20,reduction = "pca", verbose = FALSE)
```

```{r}
DimPlot(GSE128889_Filtered,group.by="DF.classifications_0.25_0.21_838")
```




```{r}
GSE128889_Filtered
GSE128889_Filtered <- subset(GSE128889_Filtered, subset = DF.classifications_0.25_0.21_838 == "Singlet")
GSE128889_Filtered 
```

from 18433 genes across 10913 cells

After DB removal 18433 features across 10075 samples 



```{r}
head(GSE128889_Filtered)
```

```{r}
GSE128889_Filtered$RNA_snn_res.0.8<- NULL
GSE128889_Filtered$pANN_0.25_0.21_838<- NULL
GSE128889_Filtered$DF.classifications_0.25_0.21_838<- NULL
GSE128889_Filtered$Study <- "GSE128889"
GSE128889_Filtered$Condition <- "Healthy Obese"
GSE128889_Filtered$sex <- "female"
GSE128889_Filtered$technology <- "Chromium-v2 single cell"
```

```{r}
head(GSE128889_Filtered)

```

```{r}
saveRDS(GSE128889_Filtered,"New_GSE128889.RDS")
```

