---
title: "New_Tabula_Spaiens"
output: html_notebook
---



#Tbula Sapiens Raw data downloaded from AWS
```{r}
set.seed(001)
```

```{r, results='hide'}
library(dplyr)
library(Seurat)
library(patchwork)
library(DoubletFinder)
```
```{r}
setwd("/home/lucamannino/R_Projects/scAtlas/Individual_Studies/Tabula_Sapiens")
```

```{r, results='hide'}
# Load the PBMC dataset
TSP10 <- Read10X(data.dir = "./TSP10/")
TSP10
# Initialize the Seurat object with the raw (non-normalized data).
TSP10 <- CreateSeuratObject(counts = TSP10, project = "TSP10", min.cells = 3, min.features = 200)
TSP10
```
10983 samples with 58870 genes obtained from cell ranger
After removing cell with less than 200 gene counts:
28171 features across 10981 samples within 1 assay 



```{r, results='hide'}
TSP14 <- Read10X(data.dir = "./TSP14/")
# Initialize the Seurat object with the raw (non-normalized data).
TSP14
TSP14 <- CreateSeuratObject(counts = TSP14, project = "TSP14", min.cells = 3, min.features = 200)
TSP14
```
58857 genes with 6794805 samples obtained from cell ranger
After removing cell with less than 200 gene counts and removing genes expressed in less than 3 cells:
26890 features across 8278 samples within 1 assay 






```{r}
TabulaSapiens <- c(TSP10,TSP14)
names(TabulaSapiens) <- c("TSP10","TSP14")
```

```{r}

TabulaSapiens$TSP10[["mt.percent"]] <-PercentageFeatureSet(TabulaSapiens$TSP10, pattern = "^MT-")
TabulaSapiens$TSP14[["mt.percent"]] <-PercentageFeatureSet(TabulaSapiens$TSP14, pattern = "^MT-")




```
```{r}
lapply(TabulaSapiens, function(x)  {VlnPlot(x, features = c("nFeature_RNA", "nCount_RNA", "mt.percent"), ncol = 3)})
```


```{r}




max.mito.thr_TSP10 <- median(TabulaSapiens$TSP10@meta.data$mt.percent) + 4*mad(TabulaSapiens$TSP10@meta.data$mt.percent)
max.mito.thr_TSP14 <- median(TabulaSapiens$TSP14@meta.data$mt.percent) + 4*mad(TabulaSapiens$TSP14@meta.data$mt.percent)
max.mito.thr_TSP10
max.mito.thr_TSP14
```
#the mitochondria percentages are way too high will need a stricter amount


```{r}
library(ggExtra)
library(ggplot2)
p1 <- ggplot(TabulaSapiens$TSP10@meta.data, aes(x=nFeature_RNA, y=mt.percent)) +
      geom_point() +
      geom_hline(aes(yintercept = 20), colour = "red", linetype = 2) #+
#      annotate(geom = "text", label = paste0(as.numeric(table(TabulaSapiens$TSP10@meta.data$mt.percent > max.mito.thr_TSP10)))," cells removed\n",                                        as.numeric(table(TabulaSapiens$TSP10@meta.data$mt.percent > max.mito.thr_TSP10)," cells remain"), x = 6000, y = 0.1)

ggMarginal(p1, type = "histogram", fill="lightgrey", bins=100) 
```

```{r}
p2 <- ggplot(TabulaSapiens$TSP14@meta.data, aes(x=nFeature_RNA, y=mt.percent)) +
      geom_point() +
      geom_hline(aes(yintercept = 20), colour = "red", linetype = 2) #+
#      annotate(geom = "text", label = paste0(as.numeric(table(TabulaSapiens$TSP10@meta.data$mt.percent > max.mito.thr_TSP10)))," cells removed\n",                                        as.numeric(table(TabulaSapiens$TSP10@meta.data$mt.percent > max.mito.thr_TSP10)," cells remain"), x = 6000, y = 0.1)

ggMarginal(p2, type = "histogram", fill="lightgrey", bins=100) 
```


```{r}

lapply(TabulaSapiens, function(x)  {FeatureScatter(x, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")})
lapply(TabulaSapiens, function(x)  {ggMarginal(ggplot(x@meta.data, aes(x=nFeature_RNA, y=nCount_RNA))+ geom_point() , type = "histogram", fill="lightgrey", bins=100)})
```
similar threshold was chosen for all samples

```{r}
TabulaSapiens$TSP10
TabulaSapiens$TSP10 <- subset(TabulaSapiens$TSP10, subset = nFeature_RNA < 7500 & mt.percent < 20  & nCount_RNA < 50000 )
TabulaSapiens$TSP14
TabulaSapiens$TSP14 <- subset(TabulaSapiens$TSP14, subset = nFeature_RNA < 7500 & mt.percent < 20  & nCount_RNA < 50000 )


```

```{r, results='hide'}
TabulaSapiens$TSP10
TabulaSapiens$TSP14


```

TSP10 from:28171 features across 10981 samples to: 28171 features across 6628 samples
TSP14 from 26890 features across 8278 samples to: 26890 features across 5032 samples






```{r}

lin_mod <- lm(log(TabulaSapiens$TSP10@meta.data$nFeature_RNA) ~ log(TabulaSapiens$TSP10@meta.data$nCount_RNA))
sub <- lin_mod$model[lin_mod$residuals < -0.5,]
plot(log(TabulaSapiens$TSP10@meta.data$nCount_RNA),log(TabulaSapiens$TSP10@meta.data$nFeature_RNA), xlab ="log(nCount_RNA)",ylab="log(nFeature_RNA)",pch=20,col="grey") +
abline(lin_mod,col="red",lwd=3) +
points(sub[,2], sub[,1],pch=20,col="blue")
tokeep <- as.numeric(names(lin_mod$residuals[lin_mod$residuals >= -0.5]))
TabulaSapiens$TSP10 <- subset(TabulaSapiens$TSP10, cells = tokeep)
```
```{r}
lin_mod <- lm(log(TabulaSapiens$TSP14@meta.data$nFeature_RNA) ~ log(TabulaSapiens$TSP14@meta.data$nCount_RNA))
sub <- lin_mod$model[lin_mod$residuals < -0.5,]
plot(log(TabulaSapiens$TSP14@meta.data$nCount_RNA),log(TabulaSapiens$TSP14@meta.data$nFeature_RNA), xlab ="log(nCount_RNA)",ylab="log(nFeature_RNA)",pch=20,col="grey") +
abline(lin_mod,col="red",lwd=3) +
points(sub[,2], sub[,1],pch=20,col="blue")
tokeep <- as.numeric(names(lin_mod$residuals[lin_mod$residuals >= -0.5]))
TabulaSapiens$TSP14 <- subset(TabulaSapiens$TSP14, cells = tokeep)
```


```{r, results='hide'}
TabulaSapiens$TSP10
TabulaSapiens$TSP14

```

TSP10 28171 features across 6612 samples
TSP14 26890 features across 4992 samples 




Nest step remove doublets with doublet finder
estimated the number of doublets by multiplying the number of total cells sequenced by 0.08 per 1000 cells
First step is estimate percentage of doublets:
```{r}
TSP10DoubletPerc <- 0.008*(10981/1000)
TSP10DoubletPerc
```
DB proportion:0.087848

```{r}
TabulaSapiens$TSP10 <- TabulaSapiens$TSP10 %>% NormalizeData(normalization.method = "LogNormalize", scale.factor = 10000) %>% FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% ScaleData(vars.to.regress = c("nCount_RNA", "nFeature_RNA")) %>% RunPCA(npcs = 100)

ElbowPlot(TabulaSapiens$TSP10, ndims = 50)

sweep.res.list_TabulaSapiensTSP10 <- paramSweep_v3(TabulaSapiens$TSP10, PCs = 1:30, sct = FALSE)
sweep.stats_TabulaSapiensTSP10 <- summarizeSweep(sweep.res.list_TabulaSapiensTSP10, GT = FALSE)
bcmvn_TabulaSapiensTSP10 <- find.pK(sweep.stats_TabulaSapiensTSP10)

TabulaSapiens$TSP10 <- FindNeighbors(TabulaSapiens$TSP10, dims = 1:30, reduction = "pca", features = VariableFeatures(TabulaSapiens$TSP10))
TabulaSapiens$TSP10 <- FindClusters(TabulaSapiens$TSP10)


homotypic.prop_TabulaSapiensTSP10 <- modelHomotypic(TabulaSapiens$TSP10@meta.data$seurat_clusters)   
nExp_poi_TabulaSapiensTSP10 <- round(TSP10DoubletPerc*nrow(TabulaSapiens$TSP10@meta.data))
nExp_poi.adj_TabulaSapiensTSP10 <- round(nExp_poi_TabulaSapiensTSP10*(1-homotypic.prop_TabulaSapiensTSP10))


pK_TabulaSapiensTSP10 = as.numeric(as.character(bcmvn_TabulaSapiensTSP10$pK))
BCmetric_TabulaSapiensTSP10 = bcmvn_TabulaSapiensTSP10$BCmetric
pK_choose_TabulaSapiensTSP10 = pK_TabulaSapiensTSP10[which(BCmetric_TabulaSapiensTSP10 %in% max(BCmetric_TabulaSapiensTSP10))]

par(mar=c(5,4,4,8)+1,cex.main=1.2,font.main=2)
plot(x = pK_TabulaSapiensTSP10, y = BCmetric_TabulaSapiensTSP10, pch = 16,type="b",
col = "blue",lty=1)
abline(v=pK_choose_TabulaSapiensTSP10,lwd=2,col='red',lty=2)
title("The BCmvn distributions")
text(pK_choose_TabulaSapiensTSP10,max(BCmetric_TabulaSapiensTSP10),as.character(pK_choose_TabulaSapiensTSP10),pos = 4,col = "red")
```

use appropiate pK value as estimated above
```{r}
TabulaSapiens$TSP10 <- doubletFinder_v3(TabulaSapiens$TSP10, PCs = 1:30, pN = 0.25, pK = pK_choose_TabulaSapiensTSP10, nExp = nExp_poi.adj_TabulaSapiensTSP10, reuse.pANN = FALSE, sct = FALSE)

#
```
```{r}
TabulaSapiens$TSP10 <- RunUMAP(TabulaSapiens$TSP10, dims = 1:20,reduction = "pca", verbose = FALSE)
```

```{r}
DimPlot(TabulaSapiens$TSP10,group.by="DF.classifications_0.25_0.07_543")
```




```{r}
TabulaSapiens$TSP10
TabulaSapiens$TSP10 <- subset(TabulaSapiens$TSP10, subset = DF.classifications_0.25_0.07_543 == "Singlet")
TabulaSapiens$TSP10 
```
Old Object:
28171 features across 6612 samples
Doublet removed new object:
28171 features across 6069 samples

```{r}
TSP14
TabulaSapiens$TSP14
```

```{r}
TSP14Doubletpercentage <- (8278/1000)*0.008
TSP14Doubletpercentage
```
0.066224 doublet proportion
```{r}
TabulaSapiens$TSP14 <- TabulaSapiens$TSP14 %>% NormalizeData(normalization.method = "LogNormalize", scale.factor = 10000) %>% FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% ScaleData(vars.to.regress = c("nCount_RNA", "nFeature_RNA")) %>% RunPCA(npcs = 100)

ElbowPlot(TabulaSapiens$TSP14, ndims = 50)

sweep.res.list_TabulaSapiensTSP14 <- paramSweep_v3(TabulaSapiens$TSP14, PCs = 1:30, sct = FALSE)
sweep.stats_TabulaSapiensTSP14 <- summarizeSweep(sweep.res.list_TabulaSapiensTSP14, GT = FALSE)
bcmvn_TabulaSapiensTSP14 <- find.pK(sweep.stats_TabulaSapiensTSP14)

TabulaSapiens$TSP14 <- FindNeighbors(TabulaSapiens$TSP14, dims = 1:30, reduction = "pca", features = VariableFeatures(TabulaSapiens$TSP14))
TabulaSapiens$TSP14 <- FindClusters(TabulaSapiens$TSP14)


homotypic.prop_TabulaSapiensTSP14 <- modelHomotypic(TabulaSapiens$TSP14@meta.data$seurat_clusters)   
nExp_poi_TabulaSapiensTSP14 <- round(TSP14Doubletpercentage*nrow(TabulaSapiens$TSP14@meta.data))
nExp_poi.adj_TabulaSapiensTSP14 <- round(nExp_poi_TabulaSapiensTSP14*(1-homotypic.prop_TabulaSapiensTSP14))


pK_TabulaSapiensTSP14 = as.numeric(as.character(bcmvn_TabulaSapiensTSP14$pK))
BCmetric_TabulaSapiensTSP14 = bcmvn_TabulaSapiensTSP14$BCmetric
pK_choose_TabulaSapiensTSP14 = pK_TabulaSapiensTSP14[which(BCmetric_TabulaSapiensTSP14 %in% max(BCmetric_TabulaSapiensTSP14))]

par(mar=c(5,4,4,8)+1,cex.main=1.2,font.main=2)
plot(x = pK_TabulaSapiensTSP14, y = BCmetric_TabulaSapiensTSP14, pch = 16,type="b",
col = "blue",lty=1)
abline(v=pK_choose_TabulaSapiensTSP14,lwd=2,col='red',lty=2)
title("The BCmvn distributions")
text(pK_choose_TabulaSapiensTSP14,max(BCmetric_TabulaSapiensTSP14),as.character(pK_choose_TabulaSapiensTSP14),pos = 4,col = "red")

TabulaSapiens$TSP14 <- doubletFinder_v3(TabulaSapiens$TSP14, PCs = 1:30, pN = 0.25, pK = pK_choose_TabulaSapiensTSP14, nExp = nExp_poi.adj_TabulaSapiensTSP14, reuse.pANN = FALSE, sct = FALSE)
```
```{r}
TabulaSapiens$TSP14 <- RunUMAP(TabulaSapiens$TSP14,dims = 1:20)
```


```{r}
DimPlot(TabulaSapiens$TSP14, group.by = "DF.classifications_0.25_0.14_299")
```


```{r}
#
TabulaSapiens$TSP14
TabulaSapiens$TSP14 <- subset(TabulaSapiens$TSP14, subset = DF.classifications_0.25_0.14_299 == "Singlet")

TabulaSapiens$TSP14
```

Before Doublet Removal
26890 features across 4992 samples
After Doublet Removal:
26890 features across 4693 samples


```{r}
TabulaSapiens$TSP10
TabulaSapiens$TSP14
```



```{r}
TabulaSapiens$TSP10$Study <- "TabulaSapiens"
TabulaSapiens$TSP14$Study <- "TabulaSapiens"

```
```{r}
TabulaSapiens_merged <- merge(TabulaSapiens$TSP10,TabulaSapiens$TSP14)
```

```{r}
head(TabulaSapiens_merged)
```


```{r}
TabulaSapiens_merged$RNA_snn_res.0.8 <- NULL
TabulaSapiens_merged$seurat_clusters <- NULL
TabulaSapiens_merged$pANN_0.25_0.07_543 <- NULL
TabulaSapiens_merged$DF.classifications_0.25_0.07_543 <- NULL
TabulaSapiens_merged$pANN_0.25_0.14_299 <- NULL
TabulaSapiens_merged$DF.classifications_0.25_0.14_299 <- NULL


TabulaSapiens_merged$Study <- "TabulaSapiens"
TabulaSapiens_merged$technology <- "Chromium-v3 single cell"
TabulaSapiens_merged$sex <- "male"
```

10 = healthy obese
14 = Lean
```{r}
TabulaSapiens_merged$Condition<-as.character(TabulaSapiens_merged$orig.ident)

TabulaSapiens_merged$Condition <- replace(TabulaSapiens_merged$Condition, TabulaSapiens_merged$Condition %in% "TSP10", "Healthy Obese")
TabulaSapiens_merged$Condition <- replace(TabulaSapiens_merged$Condition, TabulaSapiens_merged$Condition %in% "TSP14", "Lean")
```


```{r}
saveRDS(TabulaSapiens_merged,file="New_TabulaSapiens_merged.RDS")
```



